Effects

### 2020-01-02 doc/Effects.Rmd
Thinking about Effects and such

```{r setup, include=FALSE}
setwd("C:/Users/alpeterson7/Documents/MLH1repo/")
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

library(knitr)
library(ggplot2)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)

@@

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")

#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")

```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}

#*age effects* (male, maternal age)

#*litter effect*

#*mouse room effect* (Charmony, MSC)

#*quantification repeatability*

#*spread batch and staining effects*

#*power analysis on number of cells*

#*time of year effect (on spread)*

#*dominant effect* (F1s)

@@

# Quality Effects

below is a chunk to looking at histogram/distribution and means for the MLH1 data across quality bins

```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
#remake the pooled MLH1 histgoram plot from above but facet by quality

@@

mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)
age.effect.plot.mm

ff <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "female",]
age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female quality and mean CO")+facet_wrap(~subsp)

@@

The original plots weren't very effective, so I removed them. I made

## Quality effect on SC

@@

```{r SC.hitograms, echo=FALSE}

@@

length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells

@@

@@

```{r total.sc, echo=FALSE}
#check out effects of month dissected on SC

#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating

#from total SC

#merge
ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")

ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")

@@

library(lubridate)

#ADD month 
ll$month.2 <- month(ll$euth.date.2)

#mouse level table

ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)

ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain.y) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex.y)

>>>>

```{r setup, include=FALSE}
#setwd("C:/Users/alpeterson7/Documents/MLH1repo/") #this wd is different than where the file is ... this kinda causes errors
#setwd("C:/Users/April/Desktop/MLH1repo")
#load(file="data/MLH1/MLH1_data_setup.RData")

library(knitr)
library(ggplot2)
library(plyr)
library(lattice)
library(dplyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)
library(lubridate)#for dates / months

@@

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/total_SC_30.12.19.RData")#total SC data

#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")
#save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_12.3.19.RData")
```

@@

# Cell Quality Effects

I am testing the quality effects, since it could be a source of error and potentially biased the predicted gwRR if some mice had bad technical noise.

The the plots below, each point is a single mouse's mean MLH1 count per cell across each quality score bin.

```{r MLH1.histogram.2, echo=FALSE, include=FALSE}
#thse plots not shown
#remake the pooled MLH1 histgoram plot from above but facet by quality

@@

mm <- mouse.table.mean.MLH1.obs_qual[mouse.table.mean.MLH1.obs_qual$sex == "male",]
age.effect.plot.mm <- ggplot(mm, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Male quality and mean CO")+facet_wrap(~subsp)+
  theme(legend.position="none")

age.effect.plot.mm

ff <- mouse.table.mean.MLH1.obs_qual[(mouse.table.mean.MLH1.obs_qual$sex == "female"),]
ff <- ff[ff$strain != "CAST",]

age.effect.plot.ff <- ggplot(ff, aes(y=mean.MLH1, x=quality, color=(strain) ))+geom_point()+ 
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
  #are the lines
  ggtitle("Female quality and mean CO")+facet_wrap(~subsp)+theme(legend.position="none")

@@

These plots show that generally there is a positive relationship between good quality (1) and more MLH1 counts.

I should focus on addressing if there are differences in the relationship of quality-MLH1 counts across sexes.

## Cell Quality effect on SC

Examine image quality or other scores to test how well the python scripts segment the SC area for all the images.

@@

```{r SC.hitograms, echo=FALSE, fig.width = 4, fig.height=.1}
#try adjusting the size for thi..

@@

#length(MLH1.merge.Skel$Random.Name[MLH1.merge.Skel$skel_size > 30000])#67 large SC cells

@@

Testing the effects of mean total SC per mouse with when they were dissected.

@@

```{r total.sc, echo=FALSE}
#check out effects of month dissected on SC
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#from total SC

#merge
ll <- merge(mouse.level.skel, AP_mouse_table_w.Ages, by = "mouse")  #no
ll$euth.date.2 <- as.Date(ll$euth.date, "%m/%d/%Y")
#ncells is from the full MLH1 DF
#Ncells is from the cleaned, HQ DF

@@

ll <- subset(ll, select = -c(strain.x,strain.y, category.x, category.y, sex.x, sex.y, subsp.x, subsp.y, mouse.1, nmice) )

#ADD month number
ll$month.2 <- month(ll$euth.date.2)

ll.plot <- ggplot(ll, aes(y=mean.skel, x=euth.date.2, color=(strain) ))+geom_point()+
 scale_color_manual(values=colors_of_strains)+
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("")+facet_wrap(~sex)

ll.month.plot <- ggplot(ll, aes(y=mean.skel, x=month.2, color=(strain) ))+geom_point()+
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+
  ggtitle("")+facet_wrap(~sex)
